---
title: "Data frames with dplyr"
format: html
editor: visual
---

The exercises in this document follow along closely with [Lesson 13](https://swcarpentry.github.io/r-novice-gapminder/13-dplyr.html) of [R for Reproducible Scientific Analysis](https://swcarpentry.github.io/r-novice-gapminder/), provided by the [Software Carpentry](https://software-carpentry.org/) community under the creative commons license, [CC-BY 4.0](https://creativecommons.org/licenses/by/4.0/). For more information see [The Carpentries](https://carpentries.org/).

::: callout-tip
## Key Points

-   Use the `dplyr` package to manipulate data frames.
-   Use `select()` to choose variables from a data frame.
-   Use `filter()` to choose data based on values.
-   Use `group_by()` and `summarize()` to work with subsets of data.
-   Use `mutate()` to create new variables.
:::

```{r}
#| include: false
library(gapminder)
data(gapminder)

library(dplyr)
```

#### Challenge 1

Write a single command (which can span multiple lines and includes pipes) that will produce a data frame that has the African values for `lifeExp`, `country` and `year`, but not for other Continents. How many rows does your data frame have and why?

<!-- Challenge 1 answer begin -->

```{r}
africa<-filter(gapminder, continent =='Africa')%>%
  select(lifeExp, country, year)
# the table produces 624 rows x 3 columns 
```

this subset of `gapminder` contains `r nrow(africa)` and `r ncol(africa)` columns. This is due to amount years within the data set. There are `r length(unique(africa$year))` years worth of data for Africa. Additionally it includes `r length(unique(africa$country))` countries within Africa.

<!-- Challenge 1 answer end -->

#### Challenge 2

Calculate the average life expectancy per country. Which has the longest average life expectancy and which has the shortest average life expectancy?

<!-- Challenge 2 answer begin -->

```{r}
ave_lifeExp <-group_by(gapminder, country) %>%
  summarize(mean_lifeExp = mean(lifeExp)) %>%
  ungroup()
filter(ave_lifeExp, mean_lifeExp == max(mean_lifeExp))
filter(ave_lifeExp, mean_lifeExp == min(mean_lifeExp))

```

The country with the longest life expectancy is Iceland at a mean age of 76 years old. The country with shortest life expectancy is Sierra Leone with the average age being 36 years old.

<!-- Challenge 2 answer end -->

#### Challenge 3

Calculate the average life expectancy in 2002 of 2 randomly selected countries for each continent. Then arrange the continent names in reverse order. **Hint:** Use the `dplyr` functions `arrange()` and `slice_sample()`, they have similar syntax to other dplyr functions.

<!-- Challenge 3 answer begin -->

```{r}
set.seed(256254)
gapminder %>%
  filter( year == 2002)%>%
  group_by(continent)%>%
  slice_sample(n=2) %>%
  summarize(mean_lifeExp = mean(lifeExp))%>%
  ungroup() %>%
  arrange(desc(continent))
```

<!-- Challenge 3 answer end -->
